---
title: "Exercise 2 - Bayesian"
author: "Nina van Gerwen (1860852)"
date: "2/22/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## A: Loading the dataset

```{r}
library(haven)
library(rjags)

ex2 <- read_sav("Exercise 2 - Data.sav")
```

## B: Specifying the regression model

This is done in a separate text file.

## C: Specifying initial values

They need to be in a list of a list, initial list should be same length as the number of chains
```{r}
model.def <- jags.model(file = "Exercise 2 - Model.txt", data = ex2, n.chains = 2, inits = list(beta0 = 0, beta1 = 3, beta2 = 2.03))

update(model.def, n.iter = 1500)
```

## D: Obtaining samples from the posterior

```{r}
parameters <- c("beta0", "beta1", "beta2", "sigma")

set.seed(3)

results <- coda.samples(model = model.def, variable.names = parameters, n.iter = 10000)
summary(results)
```

## E: Assessing convergence

```{r}
## History plot per parameter
plot(results)

## Autocorrelation plots
autocorr.plot(results)

## Gelman-Rubin diagnostic plot
gelman.plot(results)
```

## F: Substantive interpretation

Visual indication of the history plot shows sound convergence. Furthermore, the autocorrelation plots also shows that the autocorrelation for all parameters converges to 0. Finally, the Gelman-Rubin diagnostic plot also shows sound convergence for every parameter.

A multiple linear regression with agreeableness and extraversion as independent variables and attitude towards pet ownership as the dependent variable shows that both independent variables 95% HDI contain zero. Hence, they are nonsignificant and not particularly useful in predicting attitude towards pet ownership.

## G: Interaction

Added an interaction term to the text file and redid the analyses.
```{r}
int.model.def <- jags.model(file = "Exercise 2 - Model.txt", data = ex2, n.chains = 2, inits = list(beta0 = 0, beta1 = 3, beta2 = 2.03, beta3 = 1.09))

update(int.model.def, n.iter = 1500)

int.parameters <- c("beta0", "beta1", "beta2", "beta3", "sigma")

set.seed(3)

int.results <- coda.samples(model = int.model.def, variable.names = int.parameters, n.iter = 10000)
summary(int.results)

plot(int.results)

autocorr.plot(int.results)

gelman.plot(int.results)
```

## Missing Data

Did not do. Make notes in class

